---
title: "Introduction to Mexican Regional Gross Domestic Product"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Mexican Regional Gross Domestic Product}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mrgdp)
```

In this user guide you can see how to use Mexican Regional Gross Domestic Product package (mrgdp). We built several functions to download economic census, filter municipal data and join several years.

The arguments for functions are: 

* year (2004, 2009, 2014 and 2019)
* state:

| state   | entitie             |
----------|---------------------|
| "ags"   | Aguascalientes      |
| "bc"    | Baja California     |
| "bcs"   | Baja California Sur |
|"camp"   | Campeche            |
| "coah"  | Coahuila            |
| "col"   | Colima              |
| "chis"  | Chiapas             |
| "chih"  | Chihuahua           |
| "cdmx"  | Ciudad de México    |
| "dgo"   | Durango             |
| "gto"   | Guanajuato          |
| "gro"   | Guerrero            |
| "hgo"   | Hidalgo             |
| "jal"   | Jalisco             |
| "mex"   | México              |
|"mich"   | Michoacán           |
| "mor"   | Morelos             |
| "nay"   | Nayarit             |
| "nl"    | Nuevo León          |
| "pue"   |  Puebla             |
| "qro"   | Queretaro           |
| "qroo"  | Quintana Roo        |
| "slp"   | San Luis Potosí     |
| "sin"   | Sinaloa             |
| "son"   | Sonora              |
| "tab"   | Tabasco             |
| "tamps" | Tamaulipas          |
| "tlax"  | Tlaxcala            |
| "ver"   | Veracruz            |
| "yuc"   | Yucatán             |
| "zac"   | Zacatecas           |
| "oax"   | Oaxaca              |

```{r}
#data <- economic_census(2004, "ags")
#names(data)
```

  [1] "entidad"    "municipio"  "codigo"     "id_estrato"
  [5] "ue"         "a111a"      "a121a"      "a131a"     
  [9] "a211a"      "a221a"      "a511a"      "a700a"     
 [13] "a800a"      "h000a"      "h000b"      "h000c"     
 [17] "h000d"      "h001a"      "h001b"      "h001c"     
 [21] "h001d"      "h010a"      "h010b"      "h010c"     
 [25] "h010d"      "h020a"      "h020b"      "h020c"     
 [29] "h020d"      "h101a"      "h101b"      "h101c"     
 [33] "h101d"      "h203a"      "h203b"      "h203c"     
 [37] "h203d"      "i000a"      "i000b"      "i000c"     
 [41] "i000d"      "i100a"      "i100b"      "i100c"     
 [45] "i100d"      "i200a"      "i200b"      "i200c"     
 [49] "i200d"      "j000a"      "j010a"      "j203a"     
 [53] "j300a"      "j400a"      "j500a"      "j600a"     
 [57] "k000a"      "k010a"      "k020a"      "k030a"     
 [61] "k042a"      "k412a"      "k050a"      "k060a"     
 [65] "k070a"      "k090a"      "k096a"      "k311a"     
 [69] "k610a"      "k620a"      "k810a"      "k820a"     
 [73] "k910a"      "k950a"      "k976a"      "m000a"     
 [77] "m010a"      "m020a"      "m030a"      "m050a"     
 [81] "m090a"      "m091a"      "m700a"      "o010a"     
 [85] "o020a"      "p000a"      "p000b"      "p000c"     
 [89] "p030a"      "p030b"      "p030c"      "p100a"     
 [93] "p100b"      "q000a"      "q000b"      "q000c"     
 [97] "q000d"      "q010a"      "q020a"      "q030a"     
[101] "q400a"      "q900a"      "a112a"      "a113a"     
[105] "a114a"      "a115a"      "a116a"      "a117a"     
[109] "a118a"      "a119a"      "a141a"      "a143a"     
[113] "a144a"      "a145a"      "a146a"      "a147a"     
[117] "a148a"      "a149a"      "a151a"      "a152a"     
[121] "a153a"      "a154a"      "a155a"      "a156a"     
[125] "a157a"      "a158a"      "a159a"      "a171a"     
[129] "a173a"      "a174a"      "a175a"      "a176a"     
[133] "a177a"      "a179a"      "a181a"      "a182a"     
[137] "a192a"      "a193a"      "a194a"      "a195a"     
[141] "a197a"      "a201a"      "a202a"      "a204a"     
[145] "a206a"      "a208a"      "a209a"      "a212a"     
[149] "a213a"      "a214a"      "a215a"      "a216a"     
[153] "a217a"      "a218a"      "a219a"      "a222a"     
[157] "a223a"      "a224a"      "a225a"      "a226a"     
[161] "a227a"      "a228a"      "a229a"      "a231a"     
[165] "a235a"      "a236a"      "a237a"      "a238a"     
[169] "a242a"      "a243a"      "a244a"      "a245a"     
[173] "a247a"      "a248a"      "a252a"      "a253a"     
[177] "a256a"      "a433a"      "a434a"      "a436a"     
[181] "a529a"      "a534a"      "a747a"      "a748a"     
[185] "a752a"      "a764a"     
Show in New Window
Description:df [6 × 3]
 
 
entidad
<chr>
municipio
<dbl>
a111a
<dbl>
9013	01	6	433.305	
9793	01	7	725.663	
10663	01	8	24.947	
11009	01	9	671.078	
11379	01	10	224.185	
11677	01	11	3980.801	
6 rows
Show in New Window
Description:df [6 × 6]
 
 
Entidad
<chr>
Muicipio
<dbl>
pbt04
<dbl>
pbt09
<dbl>
pbt14
<dbl>
6	01	6	433.305	757.077	619.236	
7	01	7	725.663	1611.430	1957.359	
8	01	8	24.947	55.888	71.758	
9	01	9	671.078	1477.156	1725.812	
10	01	10	224.185	56.259	118.744	
11	01	11	3980.801	7271.449	11322.543	
6 rows | 1-6 of 6 columns



Once economic census is downloaded, we can filter values by municipality. This action is reached by pbt_mun() function:

```{r}
#ags04 <- pbt_mun(data)
#tail(ags04)
```

 
 
entidad
<chr>
municipio
<dbl>
a111a
<dbl>
9013	01	6	433.305	
9793	01	7	725.663	
10663	01	8	24.947	
11009	01	9	671.078	
11379	01	10	224.185	
11677	01	11	3980.801	
6 rows
Show in New Window
Description:df [6 × 6]
 
 
Entidad
<chr>
Muicipio
<dbl>
pbt04
<dbl>
pbt09
<dbl>
pbt14
<dbl>
6	01	6	433.305	757.077	619.236	
7	01	7	725.663	1611.430	1957.359	
8	01	8	24.947	55.888	71.758	
9	01	9	671.078	1477.156	1725.812	
10	01	10	224.185	56.259	118.744	
11	01	11	3980.801	7271.449	11322.543	
6 rows | 1-6 of 6 columns
 
 
 as we can see, tail function shows the last six observations in data set. The last one in column "municipio" indicates the number of municipalities into the entity. In this case, Aguascalientes has eleven municipalities. 
 
If we are interested in build a data set with many years for this entity, we can run the PIBM_04_19() function: 

```{r}
#ags_04_19 <- PIBM_04_19(data, "ags")
#tail(ags_04_19)
```
 
 
Entidad
<chr>
Muicipio
<dbl>
pbt04
<dbl>
pbt09
<dbl>
pbt14
<dbl>
6	01	6	433.305	757.077	619.236	
7	01	7	725.663	1611.430	1957.359	
8	01	8	24.947	55.888	71.758	
9	01	9	671.078	1477.156	1725.812	
10	01	10	224.185	56.259	118.744	
11	01	11	3980.801	7271.449	11322.543	
6 rows | 1-6 of 6 columns
